Msprime 1.0: an efficient and extensible coalescent simulation framework

### 2021-08-04 paper.tex
Changed from overleaf

Such
imprecision is unfortunate, since there are any number of graphical encodings of
genetic ancestry, and referring to the original Griffiths papers does little to
clarify. As we see below, there are situations in which the precise information
stored in the graph is important, and conflating a high-level
discussion of genetic ancestry with this mathematical object is not
helpful. The point of technical terms is to provide clarity and
precision on what is being discussed, and it is now essentially
impossible to know what is meant by the term ``ARG'';
we therefore use the term ``genetic ancestry'' in high-level
discussions.

@@

\subsection*{Integration with forward simulators}
A unique feature of \msprime\ is its ability to simulate
genetic ancestries conditioned on an existing partial
genetic ancestry. Given a tree sequence that
is complete up until time $t$ as input
(where marginal trees may or may not have fully coalesced),
\msprime\ can efficiently obtain the segments of ancestral material
present at this time, and then run the simulation backwards in time from there.
This allows a simulated ancestry to be produced by any
number of different processes across disjoint time slices.
In practise this feature is used to ``complete''
forwards-time ancestry simulations~\citep{kelleher2018efficient}
that may have not fully coalesced. This process
(``recapitation'') can be orders of magnitude faster than
the standard approach of neutral burn-in; see
\cite{haller2018tree} for more details and examples.
This interoperability between simulators, where a partial ancestry
simulation produced by \SLiM~\citep{haller2019slim}
or \fwdpy~\citep{thornton2014cpp} can be picked up and completed
by another simulator, with complete information
retained---at scale---is unprecedented.

>>>>

Such
imprecision is unfortunate, since there are many ways to encode genetic
ancestry as a graph, and referring to the original Griffiths papers does little to
clarify. As we see below, there are situations in which the precise information
stored in the graph is important, and conflating a high-level
discussion of genetic ancestry with this mathematical object is not
helpful. The point of technical terms is to provide clarity and
precision on what is being discussed, and it is now essentially
impossible to know what is meant by the term ``ARG'';
we therefore use the term ``genetic ancestry'' in high-level
discussions.

@@

\subsection*{Integration with forward simulators}
A unique feature of \msprime\ is its ability to simulate
genetic ancestries by extending an existing partial
genetic ancestry. Given a tree sequence that
is complete up until time $t$ as input
(where marginal trees may or may not have fully coalesced),
\msprime\ can efficiently obtain the segments of ancestral material
present at this time, and then run the simulation backwards in time from there.
This allows a simulated ancestry to be produced by any
number of different processes across disjoint time slices.
In practise this feature is used to ``complete''
forwards-time ancestry simulations~\citep{kelleher2018efficient}
that may have not fully coalesced. This process
(``recapitation'') can be orders of magnitude faster than
the standard approach of neutral burn-in; see
\cite{haller2018tree} for more details and examples.
This interoperability between simulators, where a partial ancestry
simulation produced by \SLiM~\citep{haller2019slim}
or \fwdpy~\citep{thornton2014cpp} can be picked up and completed
by another simulator, with complete information
retained---at scale---is unprecedented.
